---
title: "Initial Korean covid 19"
author: "Adrian Cruz"
date: "October 10"
output:
  html_document: default
  pdf_document: default
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle
  
Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

## Methods
$H_0$= Dying is independent of sex.  
$H_A$= Dying is dependent of sex.  

Sex~Result Cat~Cat




## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + sex,data=kRdata)
barchartGC(~result + sex,data=kRdata, type="percent")

```

  The bar chart shows that men die at a higher rate than women.



  
## Numerical results

```{r}
table1 <- xtabs(~result + sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

This numerical data shows that 64% of men died and only 36% of women died.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

With these inferential results, the P-Value was 0.001, thus, we must reject the null hypothesis since the P-value is lower than 0.05. Therefore, men are 2.3 more times likely to die from Covid in Korea than women. 

## Conclusions

We analyzed the data regarding women and men death rates in Korea during covid and if sex is independent of dying. We first plotted a graph and were shown that men often tend to die more than women on average. Stastically the difference was pretty massive with men dying at a 64% rate and women only at 36%. Finally, we analyzed the P-Value to guarantee that there is a difference between sex and death. After viewing that the P-value is 0.001, which is lower than 0.05, we must reject the null hypothesis meaning that death is dependent on sex. We also know that men are 2.3 times more likely to die from covid in Korea than women.